Parsing Postfix log files

### 2007-11-15 doc/logparser-acronyms.tex
Dealt with a few TODO items, but added lots more. Added lots more references in the process.

\newacronym{IP}{Internet Protocol}{description={
    XXX ADD A DEFINITION\@.
}}

>>>>

\newacronym{IP}{Internet Protocol address}{description={
    The Internet Protocol~\cite{wikipedia-ip} is the protocol used to
    communicate between computers on the Internet.  An IP address is a
    unique address assigned to a computer on the Internet, allowing it to
    communicate with other computers on the Internet.  For more information
    see~\cite{wikipedia-ip-address}.
}}

\newacronym{PDF}{Portable Document Format}{description={
    PDF is the file format created by Adobe Systems in 1993 for document
    exchange.  It aims to be device independent, so documents should look
    the same whether viewed on screen or printed.
}}

### 2007-11-15 doc/logparser.tex
Dealt with a few TODO items, but added lots more. Added lots more references in the process.

@@

            \item New columns can be added to the tables used by the
                program, with sufficient DEFAULT clauses or a clever
                TRIGGER or two.\footnote{Please refer to an \SQL{} guide
                for explanations of these terms, e.g.~\cite{wikipedia-sql}
                XXX IMPROVE.}

@@

    \item \SQL{} is reasonably standard and many people will already be
        familiar with it; for those unfamiliar with it there are lots of
        readily available resources from which to learn.  Although every
        vendor implements a different dialect of \SQL{}, the basics are the
        same everywhere (analogous to the overall similarities and minor
        differences between British English, American English and
        Australian English).  XXX CITE SOME RESOURCES\@.

@@

This section discusses the rules used in parsing Postfix log files,
starting with rule characteristics, continuing with detecting and dealing
with overlapping rules, and the problems they can cause.  An example rule
and a line it would match are provided, plus a description of how the
fields in the rule are used when matching and performing the action (the
details of the table containing the rules have already been described in
section~\refwithpage{rule attributes},).  This section finishes with a
discussion of rule efficiency concerns, referring to the graphs in
section~\refwithpage{graphs}.

@@

The parser does not try to detect overlapping rules;\footnote{It may be
possible to parse each rule's \regex{}, generate a \DFA{} from it, and
examine every \DFA{} to determine if any overlap.  The author has not
attempted to do this: such a project by itself would probably qualify for a
PhD, and may involve solving the Halting Problem and circumventing the
Church-Turing Thesis.  XXX REFERENCES FOR HALTING PROBLEM AND
CHURCH-TURING\@.} that responsibility is left to the author of the rules.
Unintentionally overlapping rules lead to inconsistent parsing and data
extraction because the order in which rules are tried against each line is
unspecified, and the first matching rule wins.  Overlapping rules are
frequently a requirement, allowing a more specific rule to match some lines
and a more general rule to match the majority, e.g.\ separating \SMTP{}
delivery to a specific sites from \SMTP{} delivery to the rest of the
world.  The algorithm provides a facility for ordering overlapping rules:
the priority field in each rule (defaults to zero).  Rules are sorted by
priority, highest first, and then rules with the same priority are sorted
by the number of successful matches when parsing the previous log file.
Negative priorities may be useful for catchall rules.

@@

@@

@@

    \item The final part of this section discusses rule efficiency,
        covering the effects of caching compiled \regexs{} and optimal
        ordering of rules, with reference to the graphs in
        appendix~\refwithpage{graphs}.

@@

@@

NEED A BETTER TERM THAN HARD-CODED, THOUGH THAT MIGHT ARISE FROM THE
PREVIOUS TODO ITEM\@.

@@

        Lire is a general purpose log parser supporting many different
        types of log file.  Its Postfix parser extracts the following data
        from Postfix logs: \textit{The email servers' reports will show you
        the number of deliveries and the volume of email delivered by day,
        the domains from which you receive or send the most emails, the
        relays most used, etc.\/}; notably rejections are not mentioned.
        It supports multiple output formats (text, \HTML{}, PDF XXX and
        Excel 95).  Logs can be imported and the extracted data saved, and
        reports generated from it later.  Accessing this data from another
        application is undocumented; given the source code it should be
        possible, with time, to access the data.

@@

% Postfix components
\glossary{name={bounce},description={
    The definitive documentation is \url{http://www.postfix.org/bounce.8.html}.
}}

@@

\glossary{name={local},description={
    The definitive documentation is \url{http://www.postfix.org/local.8.html}.
}}

@@

\glossary{name={postdrop},description={
    The definitive documentation is \url{http://www.postfix.org/postdrop.1.html}.
}}

\glossary{name={postsuper},description={
    The definitive documentation is available at \url{http://www.postfix.org/postsuper.1.html}.
}}

@@

\glossary{name={smtp},description={
    The definitive documentation is \url{http://www.postfix.org/smtp.8.html}.
}}

@@

\glossary{name={IPv4},description={
    XXX IPv4.
}}

\glossary{name={IPv6},description={
    XXX IPv6.
}}

>>>>

XXX MOVE plot-* TO A SUBDIRECTORY\@?

@@

            \item New columns can be added to the tables used by the
                program, with sufficient DEFAULT clauses or a clever
                TRIGGER or two.\footnote{Please refer to an \SQL{} guide
                for explanations of these terms,
                e.g.~\cite{sql-for-web-nerds}}

@@

    \item \SQL{} is reasonably standard and many people will already be
        familiar with it; for those unfamiliar with it there are lots of
        readily available resources from which to learn.  Although every
        vendor implements a different dialect of \SQL{}, the basics are the
        same everywhere (analogous to the overall similarities and minor
        differences between British English, American English and
        Australian English).  A good introduction to \SQL{} can be found
        at~\cite{sql-for-web-nerds}, others
        are~\cite{w3schools-sql-tutorial, sqlcourse.com}.

@@

This section discusses the rules used in parsing Postfix log files,
starting with rule characteristics, continuing with detecting and dealing
with overlapping rules, and the problems they can cause.  An example rule
and a line it would match are provided, plus a description of how the
fields in the rule are used when matching and performing the action (the
details of the table containing the rules have already been described in
section~\refwithpage{rule attributes},).  This section continues with a
discussion of rule efficiency concerns, referring to the graphs in
section~\refwithpage{graphs}, and finishes with an explanation of how to
add or edit rules, including the algorithm used to generate a new \regex{}
from unparsed lines.

@@

The parser does not try to detect overlapping rules;\footnote{It may be
possible to parse each rule's \regex{}, generate a \DFA{} from it, and
examine every \DFA{} to determine if any overlap.  The author has not
attempted to do this: such a project by itself would probably qualify for a
PhD, and may involve solving the Halting
Problem~\cite{wikipedia-halting-problem} and circumventing the
Church-Turing Thesis~\cite{wikipedia-church-turing-thesis}.} that
responsibility is left to the author of the rules.  Unintentionally
overlapping rules lead to inconsistent parsing and data extraction because
the order in which rules are tried against each line is unspecified, and
the first matching rule wins.  Overlapping rules are frequently a
requirement, allowing a more specific rule to match some lines and a more
general rule to match the majority, e.g.\ separating \SMTP{} delivery to a
specific sites from \SMTP{} delivery to the rest of the world.  The
algorithm provides a facility for ordering overlapping rules: the priority
field in each rule (defaults to zero).  Rules are sorted by priority,
highest first, and then rules with the same priority are sorted by the
number of successful matches when parsing the previous log file.  Negative
priorities may be useful for catchall rules.

@@

\label{example rule}

@@

\subsection{Creating new rules}

The logs produced by Postfix differ from installation to installation, as
administrators have the freedom to choose the subset of available
restrictions which suits their needs, including using different \RBL{}
services, policy servers, or custom rejection messages.  To facilitate
this, the parser's design separates parsing rules from parsing actions:
adding new actions is difficult, but adding new rules to parse new
rejection messages is trivial (and occurs much more frequently).  The
implementation provides three programs to ease the process of adding new
rules: \texttt{add-new-rule}, \texttt{edit-rule} and \texttt{logs2regexs}
XXX NEED BETTER NAMES\@.

\texttt{add-new-rule} is the simplest of the three, it just starts the
user's preferred editor with a template rule which the user modifies as
required, and then inserts the new rule into the database.  The various
fields have already been described in sections~\refwithpage{rule
attributes},~\refwithpage{rule characteristics} and~\refwithpage{example
rule}, and will not be repeated here.

\texttt{edit-rule} extracts a rule from the database, saves it in a human
readable form, and runs the user's preferred editor for the user to change
the rule as required.  Once the user is satisfied with the changes, the
database will be updated appropriately.

\texttt{logs2regexs}

@@

    \item The topic of rule efficiency was discussed next, covering the
        effects of caching compiled \regexs{} and optimal ordering of
        rules, with reference to the graphs in
        appendix~\refwithpage{graphs}.

    \item This section finished with an explanation of how to add or edit
        rules, including the algorithm used to generate a new \regex{} from
        unparsed lines.

@@

    \item The parser does not distinguish between logs produced by
        different sources when parsing; all results will be saved to the
        same database.  This may be viewed as an advantage, as logs from
        different sources can be combined in the same database, or it may
        be viewed as a limitation as there is no facility to distinguish
        between logs from different sources in the same database.  If the
        results of parsing logs from different sources must remain
        separate, the parser can easily be instructed to use a different
        database to store the results in.

@@

XXX NEED A BETTER TERM THAN HARD-CODED, THOUGH THAT MIGHT ARISE FROM THE
PREVIOUS TODO ITEM\@.

@@

        Lire is a general purpose log parser supporting many different
        types of log file.  Its Postfix parser extracts the following data
        from Postfix logs: \textit{The email servers' reports will show you
        the number of deliveries and the volume of email delivered by day,
        the domains from which you receive or send the most emails, the
        relays most used, etc.\/}; notably rejections are not mentioned.
        It supports multiple output formats (text, \HTML{}, \PDF{} and
        Excel 95).  Logs can be imported and the extracted data saved, and
        reports generated from it later.  Accessing this data from another
        application is undocumented; given the source code it should be
        possible, with time, to access the data.

@@

% Postfix components
\glossary{name={bounce},description={
    XXX The definitive documentation is \url{http://www.postfix.org/bounce.8.html}.
}}

@@

\glossary{name={local},description={
    XXX The definitive documentation is \url{http://www.postfix.org/local.8.html}.
}}

@@

\glossary{name={postdrop},description={
    XXX The definitive documentation is \url{http://www.postfix.org/postdrop.1.html}.
}}

\glossary{name={postsuper},description={
    XXX The definitive documentation is available at \url{http://www.postfix.org/postsuper.1.html}.
}}

@@

\glossary{name={smtp},description={
    XXX The definitive documentation is \url{http://www.postfix.org/smtp.8.html}.
}}

@@

\glossary{name={IPv4},description={
    Internet Protocol~\cite{wikipedia-ipv4} version 4 is the fourth version
    of the Internet Protocol used to interconnect computers on the
    Internet.  It is the first widely deployed version of IP, and has been
    in use for over 25 years.
}}

\glossary{name={IPv6},description={
    Internet Protocol~\cite{wikipedia-ipv6} version 6 is the latest version
    of the Internet Protocol used to interconnect computers on the
    Internet.  It is the successor to IPv4, bringing with it a greatly
    expanded address space, allowing many more computers to use the
    Internet simultaneously.  IPv4 and IPv6 will coexist for many years to
    come as existing networks transition from the former to the latter.
}}
